Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-05 paper.tex
limitation

\scriptsize  

@@

void top(unsigned int clk, 
   unsigned int a) {

@@

  u1.d = c;  }

@@

%This interaction pattern exhibits producer-consumer 
%relationship, which is also evident in case of our UART example in 
%Figure~\ref{fig:interleaving1}.  
%
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.
%
Figure~\ref{fig:interleaving1} illustrates the producer-consumer relationship 
in UART IP. The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$.  
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%
Table~\ref{transaction} shows the FW transactions $(\mathcal{FT})$ and its 
interacting HW transactions $(\mathcal{HT})$ from Figure~\ref{fig:interleaving1}. 
The description above the arrows 
in column 1 shows the events in $\mathcal{FT}$ and the letter below the
arrows denotes the event type, producer $(\mathcal{P})$ or consumer 
$(\mathcal{C})$. Arrows with empty description below denotes reading of interface
variables or local variables in $\mathcal{FT}$.  Arrows with no description
simply denotes control-flow with no events.  The entires in the right column of
Table~\ref{transaction} shows the corresponding transaction in HW modules. We
simply denote a transaction in HW module with the name of the module,
transmitter module in UART ($uart\_tx$), receiver module in UART ($uart\_rx$), or
the top UART module ($uart\_top$).
%
\begin{table}
\begin{center}
{\scriptsize
\begin{tabular}{|c|c|c|}
\hline
  Firmware & Hardware  \\
  Transaction ($\mathcal{ST}$) & Transaction ($\mathcal{HT}$) \\ 
\cline{1-2}
   $s_1 \xrightarrow[]{irq_o} s_2 \xrightarrow[\mathcal{C}]{inb()} s_4
   \xrightarrow[]{status == 0x0c} s_6 \xrightarrow[\mathcal{P}]{outb()} s_9 \rightarrow
   s_{11} \rightarrow s_{12}$  & $s_2 \xrightarrow[\mathcal{C}]{inb()} s_4\;\; uart\_{rx}$
   \\ 
   & $s_6 \xrightarrow[]{outb()} s_9\;\; uart\_{tx}$ \\  \hline

@@

   $s_1 \xrightarrow[]{\neg irq_o} s_3 \xrightarrow[]{wb\_{idle()}}
   s_5 \xrightarrow[]{wb\_{idle()}} s_8 \rightarrow
   s_{12}$ & $s_3 \xrightarrow[]{wb\_{idle()}} s_5\;\; uart\_{top}$ \\ 
   & $s_5 \xrightarrow[]{wb\_{idle()}} s_8\;\; uart\_{top}$ \\  \hline
\end{tabular}
}
\end{center}
\caption{\scriptsize Firmware Transactions and its interacting HW transaction}
\label{transaction}
\end{table}  
%

\textbf{Step 3: Co-verification Model}
An unified co-verification model is constructed by instrumenting  
the software netlist model of HW RTL inside the FW thread, thereby 
generating a sequential single-threaded co-verification model in C.  
%
We now formally define a our co-verification model, denoted by 
$\mathcal{M}_{seq}$.\rmcmt{refer moduleName} as software netlist.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \\ 
%

@@

{\scriptsize

@@

\caption{\scriptsize Design statistics for UART IP and SoC Design}

@@

\begin{table*}
\begin{center}
{
\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{3}{c|}{Monolithic} & \multicolumn{5}{c|}{Path-based} \\ 
\cline{3-10}
 Scenario &  & \multicolumn{2}{c|}{HW-CBMC} & CBMC & Klee & Symbiotic & \multicolumn{3}{c|}{VerifOx} \\ 
\cline{3-10}
      &       &  Bit-level & Word-level & $\mathcal{PN}$ & $\mathcal{PN}$ & $\mathcal{PN}$ & Total/Feasible & \%-age & SAT \\
\cline{1-7}
      &       &   Time     &   Time      & Time  &  Time & Time & Paths & Pruning & Time \\
\cline{1-10}      
\multicolumn{10}{|c|}{non-deterministic data but deterministic control (Scenario A)} \\ \hline
transmit (32) & 250 & 15.02 & 15.78 & 108.25 & \textbf{0}.\textbf{98} & 29.36 & 247104/224 & 99.90 & 1.13 \\ 
transmit (64) & 500 & 23.87 & 23.29 & 206.89 & 1.68 & 29.90 & 247104/324 & 99.86 & \textbf{1}.\textbf{61}  \\ \hline 
\multicolumn{10}{|c|}{non-deterministic data and non-deterministic control
(Scenario B)} \\ \hline
trans\_intr (32) & 250 & 14.86 & 15.98 & 69.19 & \textbf{1}.\textbf{15} & 29.53 & 247104/295 & 99.88 & 1.49 \\
trans\_intr (64) & 500 & 24.14 & 26.42 & 117.73 & \textbf{1}.\textbf{27} & 29.58 & 247104/362 & 99.85 & 1.81 \\ \hline
\multicolumn{10}{|c|}{non-deterministic data and non-deterministic control
(scenario C)} \\ \hline
loopback (8)  & 230 & 52.06 & 55.71 & 1289.85 & 10.76 & 29.01 & 247104/354 & 99.85 & \textbf{3}.\textbf{95} \\ 
loopback (16) & 500 & 122.12 & 147.65 & 5410.18 & 32.24 & 29.04 & 247104/690 & 99.72 & \textbf{12}.\textbf{89} \\ 
loopback (32) & 650 & 170.62 & 192.35 & TO & 112.32 & 28.27 & 247104/1282 & 99.48 & \textbf{21}.\textbf{85}   \\ 
loopback (64) & 1300 & 409.71 & 445.66 & TO & 494.25 & 56.71 & 247104/2566 & 98.96 & \textbf{62}.\textbf{31}  \\ \hline
\multicolumn{10}{|c|}{detecting data-path bugs in transmission mode w/o interrupt} \\ \hline
transmit (64) & 520 & 28.43 & 28.14 & 186.99 & \textbf{0}.\textbf{96} & 26.67 &
247104/324 & 99.86 & 1.12 \\ \hline
\multicolumn{10}{|c|}{detecting control bugs with interrupt enabled} \\ \hline
transmit (64) & 520 & 31.35 & 29.70 & 113.57 & 1.18 & 27.13 & 247104/362 & 99.85 & \textbf{1}.\textbf{05} \\ \hline
\multicolumn{10}{|c|}{detecting control bugs in loopback mode} \\ \hline
loopback (64) & 1300 & 443.15 & 427.68 & TO & 492.59 & 54.23 & 247104/2566 &
98.96 & \textbf{62}.\textbf{34} \\ \hline
\end{tabular}
\end{scriptsize}
}
\end{center}
\vspace{-1.3mm}
\caption{Bounded safety proof times for HW/SW co-verification of UART 
\label{table:safe}}
\end{table*}

@@

\begin{table*}
\begin{center}
{
\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{3}{c|}{Monolithic} & \multicolumn{3}{c|}{Path-based} \\ 
\cline{3-8}
 Scenario &  & \multicolumn{2}{c|}{HW-CBMC} & CBMC & Klee & Symbiotic & VerifOx \\ 
\cline{3-8}
      &       &  Bit-level & Word-level & $\mathcal{PN}$ & $\mathcal{PN}$ & $\mathcal{PN}$ & $\mathcal{PN}$ \\
\cline{1-8}
      &       &   Time     &   Time      & Time  &  Time & Time & Time \\
\cline{1-8}      
\multicolumn{8}{|c|}{non-deterministic data and non-deterministic control (safe)} \\ \hline
data\_transfer & 20 & 86.92 & 86.23 & 104.25 & 25.98 & 29.36 &
\textbf{17}.\textbf{42} \\ 
\hline
\multicolumn{8}{|c|}{non-deterministic data and non-deterministic control
(unsafe)} \\ \hline
data\_transfer & 20 & 92.63 & 91.28 & 95.76 & 5.68 & 28.14 &
\textbf{1}.\textbf{27} \\ 
\hline
\end{tabular}
\end{scriptsize}
}
\end{center}
\vspace{-1.3mm}
\caption{Bounded safety proof times for HW/SW co-verification of a SoC 
\label{table:SoC}}
\end{table*}

@@

% ===============================================================================
\section{Limitations of Proposed Approach}
% ===============================================================================
The primary motivation for constructing a sequential single threaded unified 
co-verification model is to avoid enumerating exponential number of interleavings 
between the HW and FW threads.  We have shown that this is extremely beneficial for 
co-designs that exhibits producer-consumer relationship.  To this end, we construct 
a sequential co-verification model in \verifox through appropriate composition of 
HW and FW threads.  Our path-based algorithm in \verifox is able to suitably exploit 
this interaction behavior in a sequential co-verification model through symbolic 
execution by generating relevant scenario/transaction pairs.  However, this approach 
is not applicable for co-designs that exhibits true concurrency~\cite{concurrency}, 
that is, a FW its interacting HW threads does not show producer-consumer relationship.

>>>>

%\scriptsize  

@@

void top(_Bool clk, unsigned a) {

@@

  u1.d = c;  

@@

%This interaction pattern exhibits producer-consumer 
%relationship, which is also evident in case of our UART example in 
%Figure~\ref{fig:interleaving1}.  
%
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.
%
Figure~\ref{fig:interleaving1} illustrates the producer-consumer relationship 
in UART IP. The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$.  
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%
Table~\ref{transaction} shows the FW transactions $(\mathcal{FT})$ and its 
interacting HW transactions $(\mathcal{HT})$ from Figure~\ref{fig:interleaving1}. 
The description above the arrows 
in column 1 shows the events in $\mathcal{FT}$ and the letter below the
arrows denotes the event type, producer $(\mathcal{P})$ or consumer 
$(\mathcal{C})$. Arrows with empty description below denotes reading of interface
variables or local variables in $\mathcal{FT}$.  Arrows with no description
simply denotes control-flow with no events.  The entires in the right column of
Table~\ref{transaction} shows the corresponding transaction in HW modules. We
simply denote a transaction in HW module with the name of the module,
transmitter module in UART ($uart\_tx$), receiver module in UART ($uart\_rx$), or
the top UART module ($uart\_top$).
%
\begin{table}
\begin{center}
{
 %\scriptsize
\begin{tabular}{|c|c|c|}
\hline
  Firmware & Hardware  \\
  Transaction ($\mathcal{ST}$) & Transaction ($\mathcal{HT}$) \\ 
\cline{1-2}
   $s_1 \xrightarrow[]{irq_o} s_2 \xrightarrow[\mathcal{C}]{inb()} s_4
   \xrightarrow[]{status == 0x0c} s_6 \xrightarrow[\mathcal{P}]{outb()} s_9 \rightarrow
   s_{11} \rightarrow s_{12}$  & $s_2 \xrightarrow[\mathcal{C}]{inb()} s_4\;\; uart\_{rx}$
   \\ 
   & $s_6 \xrightarrow[]{outb()} s_9\;\; uart\_{tx}$ \\  \hline

@@

   $s_1 \xrightarrow[]{\neg irq_o} s_3 \xrightarrow[]{wb\_{idle()}}
   s_5 \xrightarrow[]{wb\_{idle()}} s_8 \rightarrow
   s_{12}$ & $s_3 \xrightarrow[]{wb\_{idle()}} s_5\;\; uart\_{top}$ \\ 
   & $s_5 \xrightarrow[]{wb\_{idle()}} s_8\;\; uart\_{top}$ \\  \hline
\end{tabular}
}
\end{center}
\caption{Firmware Transactions and its interacting HW transaction}
\label{transaction}
\end{table}  
%

\textbf{Step 3: Co-verification Model}
An unified co-verification model is constructed through appropriate composition of 
FW and its interacting HW threads.  This is done by instrumenting the 
software netlist model of HW RTL inside the FW thread at every invocation 
of the interface functions.  This leads to a single-threaded sequential
co-verification model in C.  
%
We now formally define a our co-verification model, denoted by 
$\mathcal{M}_{seq}$.\rmcmt{refer moduleName} as software netlist.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of the $\mathcal{SN}$. Recall that $\mathcal{SN}$ is a
software netlist representation of HW RTL models.
The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$. \rmcmt{explain instrumentation with
diagram} \\ 
%

@@

{
%\scriptsize

@@

\caption{Design statistics for UART IP and SoC Design}

@@

\begin{table*}
\begin{center}
{
%\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{3}{c|}{Monolithic} & \multicolumn{5}{c|}{Path-based} \\ 
\cline{3-10}
 Scenario &  & \multicolumn{2}{c|}{HW-CBMC} & CBMC & Klee & Symbiotic & \multicolumn{3}{c|}{VerifOx} \\ 
\cline{3-10}
      &       &  Bit-level & Word-level & $\mathcal{PN}$ & $\mathcal{PN}$ & $\mathcal{PN}$ & Total/Feasible & \%-age & SAT \\
\cline{1-7}
      &       &   Time     &   Time      & Time  &  Time & Time & Paths & Pruning & Time \\
\cline{1-10}      
\multicolumn{10}{|c|}{non-deterministic data but deterministic control (Scenario A)} \\ \hline
transmit (32) & 250 & 15.02 & 15.78 & 108.25 & \textbf{0}.\textbf{98} & 29.36 & 247104/224 & 99.90 & 1.13 \\ 
transmit (64) & 500 & 23.87 & 23.29 & 206.89 & 1.68 & 29.90 & 247104/324 & 99.86 & \textbf{1}.\textbf{61}  \\ \hline 
\multicolumn{10}{|c|}{non-deterministic data and non-deterministic control
(Scenario B)} \\ \hline
trans\_intr (32) & 250 & 14.86 & 15.98 & 69.19 & \textbf{1}.\textbf{15} & 29.53 & 247104/295 & 99.88 & 1.49 \\
trans\_intr (64) & 500 & 24.14 & 26.42 & 117.73 & \textbf{1}.\textbf{27} & 29.58 & 247104/362 & 99.85 & 1.81 \\ \hline
\multicolumn{10}{|c|}{non-deterministic data and non-deterministic control
(scenario C)} \\ \hline
loopback (8)  & 230 & 52.06 & 55.71 & 1289.85 & 10.76 & 29.01 & 247104/354 & 99.85 & \textbf{3}.\textbf{95} \\ 
loopback (16) & 500 & 122.12 & 147.65 & 5410.18 & 32.24 & 29.04 & 247104/690 & 99.72 & \textbf{12}.\textbf{89} \\ 
loopback (32) & 650 & 170.62 & 192.35 & TO & 112.32 & 28.27 & 247104/1282 & 99.48 & \textbf{21}.\textbf{85}   \\ 
loopback (64) & 1300 & 409.71 & 445.66 & TO & 494.25 & 56.71 & 247104/2566 & 98.96 & \textbf{62}.\textbf{31}  \\ \hline
\multicolumn{10}{|c|}{detecting data-path bugs in transmission mode w/o interrupt} \\ \hline
transmit (64) & 520 & 28.43 & 28.14 & 186.99 & \textbf{0}.\textbf{96} & 26.67 &
247104/324 & 99.86 & 1.12 \\ \hline
\multicolumn{10}{|c|}{detecting control bugs with interrupt enabled} \\ \hline
transmit (64) & 520 & 31.35 & 29.70 & 113.57 & 1.18 & 27.13 & 247104/362 & 99.85 & \textbf{1}.\textbf{05} \\ \hline
\multicolumn{10}{|c|}{detecting control bugs in loopback mode} \\ \hline
loopback (64) & 1300 & 443.15 & 427.68 & TO & 492.59 & 54.23 & 247104/2566 &
98.96 & \textbf{62}.\textbf{34} \\ \hline
\end{tabular}
%\end{scriptsize}
}
\end{center}
\vspace{-1.3mm}
\caption{Bounded safety proof times for HW/SW co-verification of UART 
\label{table:safe}}
\end{table*}

@@

\begin{table*}
\begin{center}
{
%\begin{scriptsize}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
  & Bound & \multicolumn{3}{c|}{Monolithic} & \multicolumn{3}{c|}{Path-based} \\ 
\cline{3-8}
 Scenario &  & \multicolumn{2}{c|}{HW-CBMC} & CBMC & Klee & Symbiotic & VerifOx \\ 
\cline{3-8}
      &       &  Bit-level & Word-level & $\mathcal{PN}$ & $\mathcal{PN}$ & $\mathcal{PN}$ & $\mathcal{PN}$ \\
\cline{1-8}
      &       &   Time     &   Time      & Time  &  Time & Time & Time \\
\cline{1-8}      
\multicolumn{8}{|c|}{non-deterministic data and non-deterministic control (safe)} \\ \hline
data\_transfer & 20 & 86.92 & 86.23 & 104.25 & 25.98 & 29.36 &
\textbf{17}.\textbf{42} \\ 
\hline
\multicolumn{8}{|c|}{non-deterministic data and non-deterministic control
(unsafe)} \\ \hline
data\_transfer & 20 & 92.63 & 91.28 & 95.76 & 5.68 & 28.14 &
\textbf{1}.\textbf{27} \\ 
\hline
\end{tabular}
%\end{scriptsize}
}
\end{center}
\vspace{-1.3mm}
\caption{Bounded safety proof times for HW/SW co-verification of a SoC 
\label{table:SoC}}
\end{table*}

@@

% ===============================================================================
\section{Limitations of Proposed Approach}
% ===============================================================================
The primary motivation for constructing a sequential single threaded unified 
co-verification model is to avoid enumerating exponential number of interleavings 
between the HW and FW threads.  We have shown that this is extremely beneficial for 
co-designs that exhibits producer-consumer relationship.  
In particular, our path-based algorithm in \verifox is able to suitably exploit 
this interaction behavior in a sequential co-verification model through symbolic 
execution which generates relevant scenario/transaction pairs.  However, this approach 
is not applicable for co-designs that exhibits true concurrency~\cite{concurrency}, 
that is, a FW its interacting HW threads does not show producer-consumer relationship.  
In this case, it is imperative to consider all possible interleavings between
these threads in some effecient manner.  However, our experience is that such 
co-designs occurs rarely in practise.
